Do we need detailed demographic data to forecast plant population responses to climate change?

### 2016-07-11 manuscript/manuscript.tex
MEE submission version

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

@@

\usepackage{lineno}
\linenumbers
\usepackage{setspace}
\doublespacing
\usepackage{todonotes}
\usepackage[document]{ragged2e}
\pagewiselinenumbers

@@

\subsubsection{Title}\label{title}

Do we need demographic data to forecast population responses to climate
change?

\subsubsection{Authors}\label{authors}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Andrew T. Tredennick (corresponding author:
  \href{mailto:atredenn@gmail.com}{\nolinkurl{atredenn@gmail.com}})
\item
  Peter B. Adler
  (\href{mailto:peter.adler@usu.edu}{\nolinkurl{peter.adler@usu.edu}})
\end{enumerate}

\subsubsection{Affiliations}\label{affiliations}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Department of Wildland Resources and the Ecology Center, 5230 Old Main
  Hill, Utah State University, Logan, Utah 84322-5230 USA
\item
  Department of Wildland Resources and the Ecology Center, 5230 Old Main
  Hill, Utah State University, Logan, Utah 84322-5230 USA
\end{enumerate}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Rapid climate change has generated growing interest in forecasts of
  future population trajectories. Traditional population models,
  typically built using detailed demographic observations from one study
  site, can address climate change impacts at one location, but are
  difficult to scale up to the landscape and regional scales relevant to
  management decisions. An alternative is to build models using
  population-level data that is much easier to collect over broad
  spatial scales than individual-level data. However, such models ignore
  the fact that climate drives population growth through its influence
  on individual performance.
\item
  We test the consequences of aggregating individual responses when
  forecasting climate change impacts on two perennial grass species
  (\emph{Pascopyrum smithii} and \emph{Sprobolis flexuosus}). We
  parameterized two population models for each species, one based on
  individual-level data (survival, growth and recruitment) and one on
  population-level data (percent cover), and compared their accuracy,
  precision, and sensitivity to climate variables. For both models we
  used Bayesian ridge regression to identify the optimal predictive
  model in terms of climate covariate strengths.
\item
  The individual-level model was more accurate and precise than the
  aggregated model when predicting out-of-sample observations. When
  comparing climate effects from both models, the population-level model
  missed important climate effects from at least one vital rate for each
  species. Increasing the sample size at the population-level would not
  necessarily reduce forecast uncertainty; the way to reduce uncertainty
  is to capture unique climate dependence of individual vital rates.
\item
  Our analysis indicates that there is no shortcut to forecasting
  climate change impacts on plant populations --- detailed demographic
  data is essential. Despite the superiority of the individual-level
  model, the forecasts it generated still were too uncertain to be
  useful for decision-makers. We need new methods to collect demographic
  data efficiently across environmental gradients in space and time.
\end{enumerate}

\emph{Key words: forecasting, climate change, grassland, integral
projection model, population model}

@@

Perhaps the greatest challenge for ecology in the 21st century is to
forecast the impacts of environmental change (Clark et al. 2001, Petchey
et al. 2015). Forecasts require sophisticated modeling approaches that
fully account for uncertainty and variability in both ecological process
and model parameters (Luo et al. 2011, but see Perretti et al. 2013 for
an argument against modeling the ecological process). The increasing
statistical sophistication of population models (Rees and Ellner 2009)
makes them promising tools for predicting the impacts of environmental
change on species persistence and abundance. But reconciling the scales
at which population models are parameterized and the scales at which
environmental changes play out remains a challenge (Clark et al. 2010,
2012, Freckleton et al. 2011, Queenborough et al. 2011). The problem is
that most population models are built using data from a single study
site because collecting those data, which involves tracking the fates of
individuals plants, is so difficult. The resulting models cannot be
applied to the landscape and regional scales relevant to decision-making
without information about how the fitted parameters respond to spatial
variation in biotic and abiotic drivers (S{æ}ther et al. 2007). The
limited spatial extent of individual-level demographic datasets
constrains our ability to use population models to address applied
questions about the consequences of climate change.

The inability of most population models to address landscape-scale
problems may explain why land managers and conservation planners have
embraced species distribution models (SDMs) (see Guisan and Thuiller
2005 for a review). SDMs typically rely on easy-to-collect
presence/absence data (but see Clark et al. 2014 for new methods) and
remotely-sensed environmental covariates that allow researchers to model
large spatial extents (e.g., Maiorano et al. 2013). Thus, it is
relatively straightforward to parameterize and project SDMs over
landscapes and regions. However, the limitations of SDMs are well known
(Pearson and Dawson 2003, Elith and Leathwick 2009, Ara{ú}jo and
Peterson 2012). Ideally, researchers would provide managers with
landscape-scale population models, combining the extent of SDMs with
information about dynamics and species abundances (Schurr et al. 2012,
Merow et al. 2014).

Aggregate measures of population status, rather than individual
performance, offer an intriguing alternative for modeling populations
(Clark and Bj{ø}rnstad 2004, Freckleton et al. 2011). Population-level
data cannot provide inference about demographic mechanisms, but might be
sufficient for modeling future population states, especially since such
data are feasible to collect across broad spatial extents (e.g.,
Queenborough et al. 2011). The choice between individual and
population-level data involves a difficult trade-off: while
individual-level data leads to more mechanistic models, population-level
data leads to models that can be applied over greater spatial and
temporal extents. An open question is how much forecasting skill is lost
when we build models based on population rather than individual-level
data.

To date, most empirical population modelers have relied on
individual-level data, with few attempts to capitalize on
population-level measures. An important exception was an effort by
Taylor and Hastings (2004) to model the population growth rate of an
invasive species to identify the best strategies for invasion control.
They used a ``density-structured'' model where the state variable is a
discrete density state rather than a continuous density measure. Such
models do not require individual-level demographic data and can
adequately describe population dynamics. Building on Taylor and Hastings
(2004), Freckleton et al. (2011) showed that density-structured models
compare well to continuous models in theory, and Queenborough et al.
(2011) provide empirical evidence that density-structured models are
capable of reproducing population dynamics at landscape spatial scales,
even if some precision is lost when compared to fully continuous models.
The appeal of density-structured approaches is clear. However, none of
these models included environmental covariates.

Addressing climate change questions with models fit to population-level
data is potentially problematic. It is individuals, not populations,
that respond to climate variables (Clark et al. 2012). Ignoring this
fact amounts to an ``ecological fallacy'', where inference about the
individual relies on statistical inference on the group (Piantadosi et
al. 1988). Population growth (or decline) is the outcome of demographic
processes such as survival, growth, and recruitment that occur at the
level of individual plants. Climate can affect each demographic process
in unique, potentially opossing, ways (Dalgleish et al. 2011). These
unique climate responses may be difficult to resolve in statistical
models based on population-level data where demographic processes are
not identifiable. If important climate effects are missed because of the
aggregation inherent in in population-level data, then population models
built with such data will make uninformative or unreliable forecasts.

Here, we compare the forecasting skill of statistical and population
models based on aggregated, population-level data with models based on
individual-level data. We used a unique demographic dataset that tracks
the fates of individual plants from four species over 14 years to build
two kinds of single-species population models, traditional models using
individual growth, survival, and recruitment data and alternative models
based on basal cover. In both models, interannual variation is
explained, in part, by climate covariates. We first quantify forecasting
skill using cross-validation. We then performed simulations to quantify
the sensitivities of species' cover to small perturbations in average
precipitation and temperature. Based on the cross-validation results,
predictions of individual level models were clearly better, but,
unfortunately, still too uncertain to inform management decisions.

@@

Our demographic data come from the Fort Keogh Livestock and Range
Research Laboratory in eastern Montana's northern mixed prairie near
Miles City, Montana, USA (\(46^{\circ}\) 19' N, \(105^{\circ}\) 48' W).
The dataset is freely available on Ecological Archives\footnote{\url{http://esapubs.org/archive/ecol/E092/143/}}
(Anderson et al. 2011) , and interested readers should refer to the
metadata for a complete description. The site is about 800 m above sea
level and mean annual precipitation (1878-2009) is 334 mm, with most
annual precipitation falling from April through September. The community
is grass-dominated and we focused on the four most abundant grass
species: \emph{Bouteloua gracilis} (BOGR), \emph{Hesperostipa comata}
(HECO), \emph{Pascopyrum smithii} (PASM), and \emph{Poa secunda} (POSE)
(Fig. 1).

From 1932 to 1945 individual plants were identified and mapped annually
in 44 1-\(\text{m}^2\) quadrats using a pantograph. The quadrats were
distributed in six pastures, each assigned a grazing treatment of light
(1.24 ha/animal unit month), moderate (0.92 ha/aum), and heavy (0.76
ha/aum) stocking rates (two pastures per treatment). In this analysis we
account for potential differences among the grazing treatments, but do
not focus on grazing\(\times\)climate interactions. The annual maps of
the quadrats were digitized and the fates of individual plants tracked
and extracted using a computer program (Lauenroth and Adler 2008, Chu et
al. 2014). Daily climate data are available for the duration of the data
collection period (1932 - 1945) from the Miles City airport, Wiley
Field, 9 km from the study site.

We modeled each grass population based on two levels of data: individual
and quadrat (Fig. 2). The individual data is the ``raw'' data. For the
quadrat-level we data we simply sum individual basal cover for each
quadrat by species. This is equivalent to a near-perfect census of
quadrat percent cover because previous analysis shows that measurement
error at the individual-level is small (Chu and Adler 2014). Based on
these two datasets we can compare population models built using
individual-level data and aggregated, quadrat-level data.

All R code and data necessary to reproduce our analysis is archived on
GitHub as release v1.0\footnote{\emph{Note to reviewers}: so that v1.0
  will be associated with the published version of the manuscript, we
  have released v0.1 to be associated with this review version.}
(\url{http://github.com/atredennick/MicroMesoForecast/releases}). That
stable release will remain static as a record of this analysis, but
subsequent versions may appear if we update this work. We have also
deposited the v1.0 release on Dryad (\emph{link here after acceptance}).

\subsubsection{Stastical models of vital
rates}\label{stastical-models-of-vital-rates}

At both levels of inference (individual and quadrat), the building
blocks of our population models are vital rate regressions. For
individual-level data, we fit regressions for survival, growth, and
recruitment for each species. At the quadrat-level, we fit a single
regression model for population growth. We describe the statistical
models separately since fitting the models required different
approaches. All models contain five climate covariates that we chose
\emph{a priori}: ``water year'' precipitation at \emph{t}-1 (lagppt);
April through June precipitation at \emph{t}-1 and \emph{t}-2 (ppt1 and
ppt2, respectively) and April through June temperature at \emph{t}-1 and
\emph{t}-2 (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t} is
the observation year. We also include interactions among same-year
climate covariates (e.g., ppt1 \(\times\) TmeansSpr1) and climate
\(\times\) size interactions. Climate \(\times\) size interactions are
for climate main effects only; we do not include interactions between
size and pairs of interacting climate effects.

We fit all models using a hierarchical Bayesian approach. The models are
fully descibed in Appendix A, so here we focus on the main process and
the model likelihood. For the likelihood models, \(\textbf{y}^X\) is
always the relevant vector of observations for vital rate \emph{X}
(\(X = S, G, R, or P\) for survival, growth, recruitment, or population
growth). For example, \(\textbf{y}^S\) is a vector of 0's and 1's
indicating whether a genet survives from \emph{t} to \emph{t+1}, or not.

@@

We used logistic regression to model survival probability (\(S\)) of
genet \(i\) from species \(j\) in quadrat group \(Q\) from time \(t\) to
\(t+1\):

\begin{align}
\text{logit}(S_{ijQ,t}) &= \gamma^{S}_{j,t} + \phi^{S}_{jQ} + \beta^{S}_{j,t}x_{ij,t} + \omega^{S}_{j}w_{ij,t} + \nu^{S}_{j}w_{ij,t}x_{ij,t} + \theta^{S}_{jk}C_{k,t} \\
y^{S}_{ijQ,t} &\sim \text{Bernoulli}(S_{ijQ,t})
\end{align}

where \(x_{ij,t}\) is the log of genet size, \(\gamma^{S}_{j,t}\) is a
year-specific intercept, \(\beta^{S}_{j,t}\) is the year-specific slope
parameter for size, \(\phi^{S}_{jQ}\) is the random effect of quadrat
group location, and \(\theta^{S}_{k}\) is the fixed parameter for the
effect of the \(k\)th climate covariate at time \(t\) (\(C_{k,t}\)).
Note that the vector of climate covariates (\textbf{C}) includes climate
variable interactions and climate\(\times\)size interactions. We include
density-dependence by estimating the effect of crowding on the focal
individual by other individuals of the same species. \(\omega\) is the
effect of crowding and \(w_{t,Q}\) is the crowding experienced by the
focal individual at time \(t\) in quadrat group \(Q\). We include a
size\(\times\)crowding interaction effect (\(\nu^{S}\)).

We modeled growth as a Gaussian process describing genet size at time
\(t+1\) as a function of size at \(t\) and climate covariates:

\begin{align}
x_{ijQ,t+1} &= \gamma^{G}_{j,t} + \phi^{G}_{jQ} + \beta^{G}_{j,t}x_{ij,t} + \omega^{G}_{j}w_{ij,t} + \nu^{S}_{j}w_{ij,t}x_{ij,t} + \theta^{G}_{jk}C_{k,t} \\
y^{G}_{ijQ,t} &\sim \text{Normal}(x_{ijQ,t+1}, \varepsilon_{ij,t})
\end{align}

where \(x\) is log genet size and all other parameters are as described
for the survival regression. We capture non-constant error variance in
growth by modeling the variance around the growth regression
(\(\varepsilon\)) as a nonlinear function of predicted genet size:

\begin{align}
\varepsilon_{ij,t} = a e^{b x_{ijQ,t+1}}
\end{align}

Our data allows us to track new recruits, but we cannot assign a
specific parent to new genets. THerefore, we model recruitment at the
quadrat level: the number of new individuals of species \(j\) in quadrat
\(q\) recruiting at time \(t+1\) as a function of quadrat ``effective
cover'' (\(A'\)) in the previous year (\(t\)). Effective cover is a
mixture of observed cover (\(A\)) in the focal quadrat (\(q\)) and the
mean cover across the entire group (\(\bar{A}\)) of \(Q\) quadrats in
which \(q\) is located:

\begin{equation}
A'_{jq,t} = p_{j}A_{jq,t} + (1-p_{j})\bar{A}_{jQ,t}
\end{equation}

where \(p\) is a mixing fraction between 0 and 1 that is estimated
within the model.

We assume the number of individuals, \(y^{R}\), recruiting at time
\(t+1\) follows a negative binomial distribution:

\begin{equation}
y^{R}_{jq,t+1} \sim \text{NegBin}(\lambda_{jq,t+1},\zeta)
\end{equation}

where \(\lambda\) is the mean intensity and \(\zeta\) is the size
parameter. We define \(\lambda\) as:

\begin{equation}
\lambda_{jq,t+1} = A'_{jq,t}e^{(\gamma^{R}_{j,t} + \phi^{R}_{jQ} + \theta^{R}_{jk}C_{k,t} + \omega^{R}\sqrt{A'_{q,t}})}
\end{equation}

where \(A'\) is effective cover (\(\text{cm}^2\)) of species \(j\) in
quadrat \(q\) and all other terms are as in the survival and growth
regressions.

>>>>

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=black,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

@@

\usepackage{lineno}
\linenumbers
\usepackage{setspace}
\usepackage{sectsty}
\doublespacing
\usepackage{todonotes}
\usepackage[document]{ragged2e}
\usepackage{times}
\usepackage{enumitem}

@@

\textsf{\LARGE{\textbf{Do we need demographic data to forecast plant population dynamics?}}}

\renewcommand*{\thefootnote}{\fnsymbol{footnote}}

\vspace{1em}

\textsf{\normalsize{\textbf{Andrew T. Tredennick\textsuperscript{1}\footnote{Corresponding author: E-mail: atredenn@gmail.com}, Mevin B. Hooten\textsuperscript{2,3,4}, and Peter B. Adler\textsuperscript{1}}}}

\vspace{1em}

\textsf{\textit{\footnotesize{\textsuperscript{1}Department of Wildland Resources and the Ecology Center, 5230 Old Main Hill, Utah State University, Logan, Utah 84322, USA; \textsuperscript{2}U.S. Geological Survey, Colorado Cooperative Fish and Wildlife Research Unit, Fort Collins, CO 80523, USA; \textsuperscript{3}Department of Fish, Wildlife, and Conservation Biology, Colorado State University, Fort Collins, CO 80523 USA}; \textsuperscript{4}Department of Statistics, Colorado State University, Fort Collins, CO 80523 USA}}

\textbf{Running head}: Demographic data and population forecasting \\
\textbf{Word count}: 6845 (including reference section) \\
\textbf{Corresponding author}: \\
Andrew Tredennick (atredenn@gmail.com) \\
Department of Wildland Resources and the Ecology Center \\
5230 Old Main Hill \\ 
Utah State University \\ 
Logan, Utah 84322, USA

Last compile: \today

\allsectionsfont{\normalfont\sffamily\bfseries}
\renewcommand*{\thefootnote}{\arabic{footnote}} \setcounter{footnote}{0}

\rule{\textwidth}{1pt}

@@

\begin{enumerate}[label=\textbf{\arabic*}]
        \item Rapid environmental change has generated growing interest in forecasts of future population trajectories. Traditional population models built with detailed demographic observations from one study site can address the impacts of environmental change at particular locations, but are difficult to scale up to the landscape and regional scales relevant to management decisions. An alternative is to build models using population-level data which are much easier to collect than individual-level data over broad spatial scales. However, it is unknown whether models built using population-level data adequately capture the effects of density-dependence and environmental forcing that are necessary to generate skillful forecasts.
        \item Here, we test the consequences of aggregating individual responses when forecasting the population states and trajectories of four perennial grass species in a semi-arid grassland in Montana, USA. We parameterized two population models for each species, one based on individual-level data (survival, growth and recruitment) and one on population-level data (percent cover), and compared their forecasting skill and forecast horizons with and without the inclusion of climate covariates. For both models we used Bayesian ridge regression to weight the influence of climate covariates for optimal prediction. 
        \item In the absence of climate effects, we found no significant difference between the forecasting skill of models based on individual-level data and models based on population-level data. Climate effects were weak, but increased forecasting skill for two species. Increases in skill with climate covariates were similar between model types.
        \item In our case study, percent cover models generated forecasts as skillful as those from a demographic model. For the goal of forecasting, models based on aggregated individual-level data may offer a practical alternative to data-intensive demographic models. Long time series of percent cover data already exist for many plant species. Modelers should exploit these data to predict the impacts of environmental change.
    \end{enumerate}

\textsf{\textbf{Key-words:}} forecasting, climate change, grassland,
integral projection model, population model, statistical regularization,
ridge regression

@@

Perhaps the greatest challenge for ecology in the 21st century is to
forecast the impacts of environmental change (Clark et al. 2001, Petchey
et al. 2015). Forecasts require sophisticated modeling approaches that
fully account for uncertainty and variability in both ecological process
and model parameters (Luo et al. 2011, but see Perretti et al. 2013).
The increasing statistical sophistication of population models (Rees and
Ellner 2009) makes them promising tools for predicting the impacts of
environmental change on species persistence and abundance. But
reconciling the scales at which population models are parameterized with
the scales at which environmental changes play out remains a challenge
(Clark et al. 2010, 2012, Freckleton et al. 2011, Queenborough et al.
2011). Most population models are built using demographic data from a
single study site because tracking the fates of individuals is so
difficult. The resulting models cannot be applied to the landscape and
regional scales relevant to decision-making without information about
how the estimated parameters respond to spatial variation in biotic and
abiotic drivers (S{æ}ther et al. 2007). The limited spatial extent of
individual-level demographic datasets constrains our ability to use
population models to address applied questions about the consequences of
climate change.

Aggregate measures of population status, rather than individual
performance, offer an intriguing alternative for modeling populations
(Clark and Bj{ø}rnstad 2004, Freckleton et al. 2011). Population-level
data cannot provide inference about demographic mechanisms, but might be
sufficient for modeling future population states, especially because
population-level data, such as plant percent cover, are feasible to
collect across broad spatial extents (e.g., Queenborough et al. 2011).
The choice between individual and population-level data involves a
difficult trade-off: while individual-level data are necessary for
mechanistic models, population-level data enable models that can be
applied over greater spatial and temporal extents. An open question is
how much forecasting skill is lost when we build models based on
population rather than individual-level data.

To date, most empirical population modelers have relied on
individual-level data, with few attempts to capitalize on
population-level measures. An important exception was an effort by
Taylor and Hastings (2004) to model the population growth rate of an
invasive species to identify the best strategies for invasion control.
They used a ``density-structured'' model where the state variable is a
discrete density state rather than a continuous density measure. Such
models do not require individual-level demographic data and can
adequately describe population dynamics. Building on Taylor and Hastings
(2004), Freckleton et al. (2011) showed that density-structured models
compare well to continuous models in theory, and Queenborough et al.
(2011) provide empirical evidence that density-structured models are
capable of reproducing population dynamics at landscape spatial scales,
even if some precision is lost when compared to fully continuous models.
However, previous tests of density-structured models have yet to assess
their ability to forecast out-of-sample observations, and they have not
included environmental covariates, which are necessary to forecast
population responses to climate change.

Addressing climate change questions with models fit to population-level
data is potentially problematic. Population growth (or decline) is the
outcome of demographic processes such as survival, growth, and
recruitment that occur at the level of individual plants. Climate can
affect each demographic process in unique, potentially opposing, ways
(Dalgleish et al. 2011). These unique climate responses may be difficult
to resolve in statistical models based on population-level data where
demographic processes are not identifiable. Futhermore, models based on
aggregated data may reflect short-term effects of one vital rate more
than others whose importance may only emerge over the long-term. For
example, a one-year change in a plant species' cover or biomass might
reflect individual growth or shrinkage, whereas the long-term trajectory
of the population might be more influenced by recruitment. The same is
true for density dependence: intraspecific density depedence may act
most strongly on vital rates, like recruitment, that are difficult to
identify from population-level data. If density dependence and/or
important climate effects are missed because of the aggregation inherent
in in population-level data, then population models built with such data
will make uninformative or unreliable forecasts.

We compared the forecasting skill of statistical and population models
based on aggregated, population-level data with the skill of models
based on individual-level data. We used a demographic dataset that
tracks the fates of individual plants from four species over 14 years to
build two kinds of single-species population models, traditional models
using individual growth, survival, and recruitment data and alternative
models based on population-level (basal cover) data. We simulated the
models to answer two questions motivated by the fact that the effects of
intraspecific competition (density dependence) and interannual weather
variability act at the level of the individual (Clark et al. 2011).
First, can population models fit using aggregated individual-level data
(percent cover) adequately capture density dependence to produce
forecasts as skillful as those from models fit to demographic data?
Second, can population models fit using aggregated data adequately
capture the influence of climate on population growth and, in turn,
produce forecasts as skillful as those from models fit to demographic
data?

@@

Our demographic data come from a northern mixed grass prairie at the
Fort Keogh Livestock and Range Research Laboratory near Miles City,
Montana, USA (\(46^{\circ}\) 19' N, \(105^{\circ}\) 48' W). The dataset
is available on Ecological Archives\footnote{\url{http://esapubs.org/archive/ecol/E092/143/}}
(Anderson et al. 2011), and interested readers should refer to the
metadata for a complete description. The site is 800 m above sea level
and mean annual precipitation (1878-2009) is 334 mm, with most annual
precipitation falling from April through September. The community is
grass-dominated, and we focused on the four most abundant grass species:
\emph{Bouteloua gracilis} (BOGR), \emph{Hesperostipa comata} (HECO),
\emph{Pascopyrum smithii} (PASM), and \emph{Poa secunda} (POSE) (Fig.
1). \emph{B. gracilis} is a warm-season perennial grass, whereas
\emph{H. comata}, \emph{P. smithii}, and \emph{Poa secunda} are
cool-season perennial grasses. The growing season begins in early spring
(typically in April) and lasts through mid-summer (typically in June).

From 1932 to 1945, individual plants were identified and mapped annually
in 44 1-\(\text{m}^2\) quadrats using a pantograph. The quadrats were
distributed among six pastures, each assigned a grazing treatment of
light (1.24 ha/animal unit month), moderate (0.92 ha/aum), and heavy
(0.76 ha/aum) stocking rates (two pastures per treatment). In this
analysis, we accounted for potential differences among the grazing
treatments, but do not focus on grazing\(\times\)climate interactions.
The annual maps of the quadrats were digitized and the fates of
individual plants tracked and extracted using a computer program
(Lauenroth and Adler 2008, Chu et al. 2014). The permanent quadrats have
not been relocated, but their distribution in six different pastures
means the data represent a broad spatial distribution for the study
area. Daily climate data are available for the duration of the data
collection period (1932 - 1945) from the Miles City airport, Wiley
Field, 9 km from the study site.

We modeled each grass population based on two levels of data: individual
and quadrat. The individual data are the ``raw'' data. For the
quadrat-level data we simply sum individual basal cover for each quadrat
by species. This is equivalent to a near-perfect census of quadrat
percent cover because measurement error at the individual-level is small
(Chu and Adler 2015). Based on these two datasets of 13 year-to-year
transitions, we can compare population models built using
individual-level data and aggregated, quadrat-level data. At the
individual level, we explicitly model three vital rates: growth,
survival, and recruitment. At the quadrat level, we model population
growth as change in percent cover of quadrats with non-zero cover in
year \emph{t} and in year \emph{t-1}, ignoring within-quadrat
extirpation and colonization events because they are very rare in our
time series (\(N=16\) and \(N=13\), respectively, across all species).
Sample sizes for each species and vital rate model are shown in Table 1.

All R code and data necessary to reproduce our analysis is archived on
GitHub as release v1.0\footnote{\emph{Note to reviewers}: so that v1.0
  will be associated with the published version of the manuscript, we
  have released v0.2 to be associated with this review version.}
(\url{http://github.com/atredennick/MicroMesoForecast/releases}). We
have also deposited the v1.0 release on Dryad (\emph{link here after
acceptance}).

\subsubsection{Statistical models of vital
rates}\label{statistical-models-of-vital-rates}

At both levels of inference (individual and quadrat), the building
blocks of our population models are vital rate regressions. For
individual-level data, we fit regressions for survival, growth, and
recruitment for each species. At the quadrat-level, we fit a single
regression model for population growth. We describe the statistical
models separately because they required different approaches. For both
model types, we fit vital rate models with and without climate
covariates. Models with climate effects contain five climate covariates
that we chose \emph{a priori} based on previous model selection efforts
using these data (Chu et al. 2016) and expert advice (Lance Vermeire,
\emph{personal communication}): ``water year'' precipitation at
\emph{t}-2 (lagppt); April through June precipitation at \emph{t}-1 and
\emph{t} (ppt1 and ppt2, respectively) and April through June
temperature at \emph{t}-1 and \emph{t} (TmeanSpr1 and TmeanSpr2,
respectively), where \emph{t}-1 to \emph{t} is the transition of
interest. We also include interactions among same-year climate
covariates (e.g., ppt1 \(\times\) TmeansSpr1), resulting in a total of
seven climate covariates.

We fit all models using a hierarchical Bayesian approach. We focus on
the main process and the model likelihood in what follows (full model
descriptions are in the Supporting Information). For the likelihood
models, \(\textbf{y}^X\) is always the relevant vector of observations
for vital rate \emph{X} (\emph{X} = \emph{S}, \emph{G}, \emph{R}, or
\emph{P}) for survival, growth, recruitment, or population growth). For
example, \(\textbf{y}^S\) is a vector of 0s and 1s indicating whether a
genet survives from \emph{t} to \emph{t+1}, or not, for all observation
years and quadrats. All model parameters are species-specific, but we
omit subscripts for species in model descriptions below to reduce visual
clutter. For brevity, we only describe models with climate covariates
included, but models without climate covariates are simply the models
described below with the climate effects removed.

@@

We used logistic regression to model the probability that genet \emph{i}
in quadrat \emph{q} survives from time \emph{t} to \emph{t}+1
(\(s_{i,q,t}\)):

\vspace{-3em}\begin{align}
y_{i,q,t}^{S} &\sim \text{Bernoulli}(s_{i,q,t}), \\
\text{logit}(s_{i,q,t}) &= \beta_{0,t} + \beta_{s,t}x_{i,q,t} + \beta_{Q,q} + \textbf{z}'_t \boldsymbol{\beta}_c + \beta_{d,1} w_{i,t} + \beta_{d,2} (x_{i,q,t}w_{i,q,t}),
\end{align}\vspace{-3em}

where \(x_{i,q,t}\) is the log of genet \emph{i} basal area at time
\emph{t}, \(\beta_{0,t}\) is a year specific intercept, \(\beta_{Q,q}\)
is the random effect of the \emph{q}th quadrat to account for spatial
location, \(\beta_{s,t}\) is the year-specific slope parameter for size,
\textbf{z} is a vector of \emph{p} climate covariates specific to year
\emph{t}, \(\boldsymbol{\beta}_c\) is a vector of fixed climate effects
of length \emph{p}, \(\beta_{d,1}\) is the effect of intraspecific
crowding experienced by the focal genet at time \emph{t}
(\(w_{i,q,t}\)), and \(\beta_{d,2}\) is a size by crowding
(\(x_{i,q,t}w_{i,q,t}\)) interaction effect.

We follow the approach of Chu and Adler (2015) to estimate crowding,
assuming that the crowding experienced by a focal genet depends on
distance to each neighbor genet and the neighbor's size, \emph{u}:

\begin{equation}
w_{i,q,t} = \sum_k e^{-\delta d_{ik,q,t}^{2}}u_{k,q,t}.
\end{equation}

In equation 3, \(w_{i,t}\) is the crowding that genet \emph{i} in year
\emph{t} experiences from conspecific neighbors in quadrat \emph{q}. The
spatial scale over which conspecific neighbors exert influence on any
genet is determined by \(\delta\). The function is applied for all
\emph{k} conspecific genets that neighbor the focal genet at time
\emph{t}, and \(d_{ik,q,t}\) is the distance between genet \emph{i} and
conspecific genet \emph{k} in quadrat \emph{q}. We use
regression-specific (survival and growth) \(\delta\) values estimated by
Chu and Adler (2015).

We modeled growth as a Gaussian process describing log genet size
(\(y_{i,q,t+1}^{G}\)) at time \(t+1\) in quadrat \emph{q} as a function
of log size at time \(t\) and climate covariates:

\vspace{-3em}\begin{align}
y_{i,q,t+1}^G &\sim \text{Normal}(\mu_{i,q,t+1}, \sigma^2_{x_{i,q,t+1}}), \\
\mu_{i,q,t+1} &= \beta_{0,t} + \beta_{s,t}x_{i,q,t} + \beta_{Q,q} + \textbf{z}'_t \boldsymbol{\beta}_c + \beta_{d,1} w_{i,q,t} + \beta_{d,2} (x_{i,q,t}w_{i,q,t}),
\end{align}\vspace{-3em}

where \(\mu_{i,q,t+1}\) is log of genet \emph{i}s predicted size at time
\emph{t} + 1, and all other parameters are as described for the survival
regression. We capture non-constant error variance in growth by modeling
the variance in the growth regression (\(\sigma^2_{x{i,q,t+1}}\)) as a
nonlinear function of predicted genet size:

\vspace{-3em}\begin{align}
\sigma^2_{x_{i,q,t+1}} = a \, \text{exp}[b \times \mu_{i,q,t+1}],
\end{align}\vspace{-3em}

where \(\mu_{i,q,t+1}\) is log of predicted genet size predicted from
the growth regression (Eq. 4), and \emph{a} and \emph{b} are constants.

Our data allows us to track new recruits, but we cannot assign a
specific parent to new genets. Therefore, we model recruitment at the
quadrat level. We assume the number of individuals, \(y^{R}_{q,t+1}\),
recruiting at time \(t+1\) in quadrat \emph{q} follows a negative
binomial distribution:
